---
title: "Lab2"
author: Grupo 9 
date: February 23, 2026 
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Parte1}
knitr::opts_chunk$set(echo = TRUE)
movies <- read.csv("data/movies_2026.csv")

summary(movies)
library(factoextra)
library(ggplot2)
library(dplyr)
library(tidyr)


movies <- movies %>%
  mutate(genres = gsub(" \\| ", "|", genres))
movies_long <- movies %>%
  separate_rows(genres, sep = "\\|")
movies_longs <- movies_long %>%
  separate_rows(genres, sep = "\\s*\\|\\s*")

set.seed(123)
sample_movies <- movies_longs[sample(nrow(movies_longs), 5000), ]

fviz_nbclust(
  scale(na.omit(sample_movies[, sapply(sample_movies, is.numeric)])),
  kmeans,
  method = "wss"
)
```

# Tipo de Variable
#VARIABLES.

#- id
#Cuantitativo, discreta, integer

#- budget
#Cuantitativo, discreta, numeric

#- genres
#Cualitativo, nominal

#- homePage
#Cualitativo, nominal

#- productionCompany
#Cualitativo, nominal

#- productionCompanyCountry
#Cualitativo, nominal

#- productionCountry
#Cualitativo, nominal

#- revenue
#Cuantitativo, discreta, numeric

#- runtime
#Cuantitativo, discreta, integer

#- video
#Cualitativo, nominal

#- director
#Cualitativo, nominal

#- actors
#Cualitativo, nominal

#- actorsPopularity
#Cualitativo, ordinal

#- actorsCharacter
#Cualitativo, nominal

#- originalTitle
#Cualitativo, nominal

#- title
#Cualitativo, nominal

#- originalLanguage
#Cualitativo, nominal

#- popularity
#Cuantitativo, continua, numeric

#- releaseDate
#Cualitativo, ordinal

#- voteAvg
#Cuantitativo, continua, numeric

#- voteCount
#Cuantitativo, discreta, integer

#- genresAmount
#Cuantitativo, discreta, integer

#- productionCoAmount
#Cuantitativo, discreta, integer

#- productionCountriesAmount
#Cuantitativo, discreta, integer

#- actorsAmount
#Cuantitativo, discreta, integer

#- castWomenAmount
#Cuantitativo, discreta, integer

#- castMenAmount
#Cuantitativo, discreta, integer

#- releaseYear
#Cuantitativo, discreta, integer
