---
title: "Lab2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Parte1}
knitr::opts_chunk$set(echo = TRUE)
movies <- read.csv("movies_2026.csv")

summary(movies)

library(ggplot2)
library(dplyr)
library(tidyr)


movies <- movies %>%
  mutate(genres = gsub(" \\| ", "|", genres))
movies_long <- movies %>%
  separate_rows(genres, sep = "\\|")
movies_longs <- movies_long %>%
  separate_rows(genres, sep = "\\s*\\|\\s*")
movies_longss <- movies_longs %>%
  filter(budget != "0")

top10_genres <- movies_longss %>%
  group_by(genres) %>%
  summarise(
    presupuesto_promedio = mean(budget, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(presupuesto_promedio)) %>%
  slice_head(n = 10)

top10_genres

ggplot(top10_genres,aes(x = reorder(genres, presupuesto_promedio),y = presupuesto_promedio)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 10 géneros con mayor presupuesto promedio",
    x = "Género",
    y = "Presupuesto promedio"
  )


estructura <- data.frame(
  Variable = names(movies),
  Tipo_R = sapply(movies, class)
)

```

# Tipo de Variable
#VARIABLES.

#- id
#Cuantitativo, discreta, integer

#- budget
#Cuantitativo, discreta, numeric

#- genres
#Cualitativo, nominal

#- homePage
#Cualitativo, nominal

#- productionCompany
#Cualitativo, nominal

#- productionCompanyCountry
#Cualitativo, nominal

#- productionCountry
#Cualitativo, nominal

#- revenue
#Cuantitativo, discreta, numeric

#- runtime
#Cuantitativo, discreta, integer

#- video
#Cualitativo, nominal

#- director
#Cualitativo, nominal

#- actors
#Cualitativo, nominal

#- actorsPopularity
#Cualitativo, ordinal

#- actorsCharacter
#Cualitativo, nominal

#- originalTitle
#Cualitativo, nominal

#- title
#Cualitativo, nominal

#- originalLanguage
#Cualitativo, nominal

#- popularity
#Cuantitativo, continua, numeric

#- releaseDate
#Cualitativo, ordinal

#- voteAvg
#Cuantitativo, continua, numeric

#- voteCount
#Cuantitativo, discreta, integer

#- genresAmount
#Cuantitativo, discreta, integer

#- productionCoAmount
#Cuantitativo, discreta, integer

#- productionCountriesAmount
#Cuantitativo, discreta, integer

#- actorsAmount
#Cuantitativo, discreta, integer

#- castWomenAmount
#Cuantitativo, discreta, integer

#- castMenAmount
#Cuantitativo, discreta, integer

#- releaseYear
#Cuantitativo, discreta, integer
